Happiness in US

### 2023-03-16 outputs/paper/Happiness in America: A Comprehensive Analysis of the Effects of Demographic Factors.qmd
add content

---
title: "Happiness in US "
subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
author: 
  - Rae Zhang
  - Faustine Fan
  - Myra Li
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Data 
## Source and Methodology
Our study is based on data collected from the US General Social Survey (GSS). The GSS is a nationally representative cross-sectional survey conducted since 1972, aimed at collecting data on contemporary American society to monitor and explain trends in opinions, attitudes, and behaviors.
One of the unique features of the GSS is that it adapts questions from earlier surveys, enabling researchers to conduct comparisons for up to 80 years. Moreover, the GSS has taken great care to maintain survey methodology as comparable as possible over time, including keeping the same sampling approach and question wording. This approach helps to minimize potential changes due to variations in methodology and supports the study of trends in public opinion in the United States over time.

@@

The survey under consideration in this report is the General Social Survey, which has been conducted since 1972 by the NORC at the University of Chicago and funded by the National Science Foundation. The survey is designed to monitor public opinion and behavior in the United States and utilizes similar sampling and questioning approaches across its various rounds to ensure consistency. The GSS includes repeating modules and topical modules that may vary across rounds, and each participant is given a subset of these modules to respond to.

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8)

@@

### Input data ###
happiness_data<-read.csv(here::here("outputs/data/happiness_data.csv"))
sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
race_data<-read.csv(here::here("outputs/data/race_data.csv"))
degree_data<-read.csv(here::here("outputs/data/degree_data.csv"))
employee_type_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))

# Results

```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-1Happiness
#| fig-cap: "Changes in happiness levels from 1972 to 2021"

@@

fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy_level)) +
  geom_line(aes(group = Happy_level, color = Happy_level), size = 1.2) +
  scale_fill_manual(values = happy_colors) +
  scale_color_manual(values = happy_colors) +
  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
       color = "Happy level")
fig1 <- fig1  +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 12,
             x = 0.5,
             y = 1.05)
```

@@

fig2 <- fig2  +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)
```

@@

fig3 <- fig3  +
  draw_label("Trends in happiness levels for different races from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)
```

@@

fig4 <- fig4  +
  draw_label("Trends in happiness levels for different degrees from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)
```

@@

fig5 <- fig5  +
  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)
```

```{r, fig.height=16}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6Marital
#| fig-cap: "Trends in happiness levels for different marital status from 1972 to 2021"

@@

fig6 <- fig6  +
  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)
```

@@

# Results

# Discussion

@@

Weaknesses and next steps should also be included.

@@

# Appendix {.unnumbered}

# Additional details

>>>>

---
title: "Happiness in US"
subtitle: "Analysis of the Effects of Demographic Factors during 1972 to 2021"
author: 
  - Rae Zhang
  - Faustine Fan
  - Myra Li
thanks: "Code and data are available at: LINK."
date: "`r Sys.time()`"
date-format: "D MMMM YYYY"
abstract: "First sentence. Second sentence. Third sentence. Fourth sentence."
format: pdf
number-sections: true
bibliography: references.bib
---

@@

# Data 
## Source and Methodology
Our study is based on data collected from the US General Social Survey (GSS). The GSS is a nationally representative cross-sectional survey conducted since 1972, aimed at collecting data on contemporary American society to monitor and explain trends in opinions, attitudes, and behaviors.

One of the unique features of the GSS is that it adapts questions from earlier surveys, enabling researchers to conduct comparisons for up to 80 years. Moreover, the GSS has taken great care to maintain survey methodology as comparable as possible over time, including keeping the same sampling approach and question wording. This approach helps to minimize potential changes due to variations in methodology and supports the study of trends in public opinion in the United States over time.

@@

## Survey Frame and Weakness
Our survey has been carefully edited to improve upon existing questions from the original questionnaire and to add new questions that better capture the happiness of people in the USA. Specifically, we have addressed potential biases in questions and expanded the options to provide a more detailed exploration of the relationship between race and happiness. Additionally, we have added questions that focus on employment situations, income, COVID-19, mental health, future life aspirations, sense of fulfilment, and attitudes towards happiness.

We believe that our survey will provide a more comprehensive and nuanced understanding of the factors that contribute to happiness in the USA. By improving the survey questions, we hope to provide policymakers and individuals with valuable insights into how to improve overall well-being and foster a happier society.

## Strength and Weakness of Dataset 
The strength of the GSS data is that the GSS uses a large and representative sample of the US population, which allows the broader population to contribute to the data set to eliminate the regional bias that might be caused in different provinces in the US. Meanwhile the GSS dat has been conducted consistently over a long period of time, allowing researchers to tracker changes especially in social attitudes. The GSS also uses standardized questions and methods, making it easier to compare data across different time periods and populations.

The GSS data also has noticeable weakness. For example, while it covers a wide rangfe of topics, it may not capture all relevant factors for a particular research question. Meanwhile, with any survey, the GSS is subject to response bias, where certain individuals may be more or less likely to participate or answer certain questions. And the data only counts the US population, certain subgroups may be underrepresented or overrepresented in a sample.

## Variables
The 6 variables we selected from GSS for this research listed as follow:

Race: the race respondent considered themselves are

Happy: the happiness levels people felt at that moment

Degree: the highest degree respondents had till that time

Wrkslf: the employment status of the respondents

Marital:  the marital status of respondents

Sex: the sex respondents considered themselves are

# Results
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE)
knitr::opts_chunk$set(warning = FALSE)
knitr::opts_chunk$set(message = FALSE)
knitr::opts_chunk$set(fig.width=6.5, fig.height=2.8)

@@

### Input data ###
happiness_data<-read.csv(here::here("outputs/data/happiness_data.csv"))
sex_data<-read.csv(here::here("outputs/data/sex_data.csv"))
race_data<-read.csv(here::here("outputs/data/race_data.csv"))
degree_data<-read.csv(here::here("outputs/data/degree_data.csv"))
employee_type_data<-read.csv(here::here("outputs/data/employee_status_data.csv"))
marital_status_data<-read.csv(here::here("outputs/data/marital_status_data.csv"))
```

## Overall Happiness
```{r}
#| message: false
#| warning: false
#| echo: false
#| label: fig-1Happiness
#| fig-cap: "Changes in happiness levels from 1972 to 2021"

@@

fig1 <- ggplot(happiness_data, aes(x = Year, y = People, fill = Happy_level)) +
  geom_line(aes(group = Happy_level, color = Happy_level), size = 1.2) +
  scale_fill_manual(values = happy_colors) +
  scale_color_manual(values = happy_colors) +
  labs(x = "Year", y = "Number of Respondents", fill = "Happy Level", 
       color = "Happy level")
fig1 <- fig1  +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 12,
             x = 0.5,
             y = 1.05)
fig1
```
Over the past five decades, @fig-1Happiness shows there has been an overall increase in happiness levels in the United States, though with significant fluctuations. Further analysis of this trend reveals that happiness is influenced by various factors, such as macroeconomic conditions, social changes, and personal circumstances.

For instance, in the early 1990s, there was a considerable rise in happiness due to political and economic changes, such as the end of the Cold War, the emergence of neoliberalism, and the rise of a new Democratic Party (@ProtectingSoldiersandMothers). In contrast, the dot-com crash of the early 2000s led to a sharp decline in happiness levels that persisted until the economic recovery brought it back up to pre-crash levels.

However, the global financial crisis and recession of 2007-2009 resulted in another decrease in happiness, as it impacted people's quality of life. More recently, the COVID-19 pandemic has had both positive and negative effects on happiness levels. While social isolation and economic uncertainty have decreased happiness, increased family and leisure time have had positive effects.

We will examine how demographic factors such as sex, race, marital status, degree, and employment status influence happiness and why as follow.

## Sex

@@

fig2 <- fig2  +
  draw_label("Trends in happiness levels for different genders from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)

fig2
```
In this analysis, we examined @fig-2Gender, which compares the proportion of different education levels for each sex group from 1972 to 2021. We observed that the distribution of overall happiness levels remains consistent over time, with "pretty happy" being the most common response for both male and female groups.

A slight difference was noted in the mid-1970s, where the female group appeared slightly happier than the male group. However, the overall trends in happiness levels for both groups align with the trends shown in @fig-1Happiness.

It is worth noting that our analysis is limited to the data presented in the figures, and further investigation may be required to understand any potential underlying factors contributing to these trends.

## Race
In @fig-3Race, we present the trends in happiness levels among different racial groups in America from 1972 to 2021. The figure shows that white people consistently report higher levels of happiness than other racial groups over the period, although this could be due to survivorship bias as the majority of respondents in the GSS dataset are white. It is important to note that all racial groups have become less happy in 2021, possibly due to the COVID-19 pandemic.

Notably, the figure shows that black people have experienced drops in happiness levels in specific years, including 1980, 2008, and 2021. These drops in happiness levels may reflect the impact of significant events or social phenomena that occurred during those years and that disproportionately affected the happiness of the black race. Further research may be needed to explore the reasons for these trends and to determine the extent to which they are influenced by factors such as racism, economic inequality, and political events.

@@

fig3 <- fig3  +
  draw_label("Trends in happiness levels for different races from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)

fig3
```

## Degree
The trends in happiness levels across different education groups from 1972 to 2021 are presented in @fig-4Degree. It can be observed that individuals with a graduate and above education experience report the lowest levels of the "not too happy" response option and have the highest proportion of respondents selecting the "very happy" option across the years. Conversely, individuals with less than a high school education have the highest proportion of respondents selecting the "not too happy" response option among the different education groups over the years.

@@

fig4 <- fig4  +
  draw_label("Trends in happiness levels for different degrees from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)

fig4
```

## Employee Type
In general, there may not be significant differences between being self-employed or being employed by someone else in terms of overall trends and distribution. However, upon closer examination of @fig-5Employee, it becomes apparent that self-employed individuals tend to report higher levels of happiness than those who are employed by others. This could be attributed to several factors, such as the flexibility of their work schedule, the ability to reduce commute time, and fewer relationship issues. Overall, these benefits may contribute to a greater sense of control over one's work and personal life, which can positively impact overall happiness levels.

@@

fig5 <- fig5  +
  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)

fig5
```

## Marital Status 
```{r, fig.height=16}
#| message: false
#| warning: false
#| echo: false
#| label: fig-6Marital
#| fig-cap: "Trends in happiness levels for different marital status from 1972 to 2021"

@@

fig6 <- fig6  +
  draw_label("Trends in happiness levels for different emoloyee status from 1972 to 2021", 
             fontface = "bold", 
             size = 16,
             x = 0.5,
             y = 1.05)

fig6
```
@fig-6Marital indicates that married individuals are more likely to report higher levels of happiness, with a greater proportion of respondents selecting "Pretty happy" and "Very happy" compared to those in other marital status categories. Interestingly, the proportion of respondents selecting "Very happy" in other groups is comparatively lower. Additionally, the distribution of happiness levels in the married group appears to be more stable and consistent over time compared to the other groups.

It is worth noting that the COVID-19 pandemic had a significant impact on people's happiness levels, and married people’s level of happiness was more stable during this time. This may be attributed to the stability and security that marriage can provide during times of uncertainty.

@@

# Discussion
## Race and Levels of Happiness 
Upon analyzing @fig-2Gender, we observed that the overall distribution of happiness levels appears to be similar across different racial groups. However, there was a significant increase in the proportions of happiness levels reported by individuals identifying as Black or belonging to other racial groups during 1970 to 1990. This suggests that race identity may potentially influence people's happiness levels within a society. Individuals from diverse racial and ethnic groups may experience varying degrees of social and economic disadvantage, discrimination, and prejudice, which can impact their overall sense of well-being and happiness, leading to a major decrease or increase in self-reported happiness levels.

Research has shown that individuals who identify as racial minorities often experience more stress and lower levels of life satisfaction than those who identify as members of the dominant racial group (@FREDERICK20221). This can be due to factors such as experiences of racism, lack of representation in positions of power, and limited access to resources and opportunities. However, it is important to note that the impact of race identity on happiness is complex and multifaceted. The relationship between race and happiness can be influenced by factors such as social support, cultural identity, and personal values. Moreover, individual differences in personality, life circumstances, and coping strategies can also play a significant role in shaping the relationship between race identity and happiness. Therefore, while race identity can be a relevant factor in determining people's level of happiness, it is important to consider the context and individual differences in any analysis of this relationship.

## Education levels and Happiness

## Marital Status and level of Happiness

## Weakness

##Next Steps

@@

@@

\newpage
# Appendix
A link to our survey exploring what will affect happiness in US can be found in the footnotes below.^[https://forms.gle/ZAoN1uFc9Gc24anh8]

```{r, echo=FALSE, out.width="100%"}
knitr::include_graphics(here::here("outputs/survey/survey_1.png"))
knitr::include_graphics(here::here("outputs/survey/survey_2.png"))
knitr::include_graphics(here::here("outputs/survey/survey_3.png"))
knitr::include_graphics(here::here("outputs/survey/survey_4.png"))
knitr::include_graphics(here::here("outputs/survey/survey_5.png"))
knitr::include_graphics(here::here("outputs/survey/survey_6.png"))
```
# Additional details
